---
title: "Exercise 7"
author: "G.K Akwaboah"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(ggplot2)
```

1. Smoking and Birth Weight
a. Test Selection and Sample Size Calculation
Two sample T-test

# Given values
```{r}
mu1 <- 3200  # non-smokers
mu2 <- 2850  # smokers
sd <- 700    # standard deviation (same for both)
alpha <- 0.05
z_alpha <- 1.96
power <- 0.80
z_beta <- 0.84 #for 80% power
beta <- 0.20 # 1 - power
effect_size <- (mu1 - mu2) / sd
```

Least sample size needed, n
```{r}
n = (((z_alpha + z_beta)*sd)/(mu1 -mu2))^2
```

b. Load Data and Perform Test
```{r}
data(birthwt)
head(birthwt)
```

# Create smoking variable: 1 = smoker, 0 = non-smoker. convert to factor
```{r}
birthwt$smoke <- factor(birthwt$smoke, labels = c("Non-Smoker", "Smoker"))
```

# Check sample sizes
```{r}
table(birthwt$smoke)
```

# Perform two-sample t-test
```{r}
t_test_result <- t.test(bwt ~ smoke, data = birthwt, var.equal = TRUE)
t_test_result
```

c. Compare Mother's Weight Distribution
# Boxplot for mother's weight (lwt)

```{r}
ggplot(birthwt, aes(x = smoke, y = lwt)) +
  geom_boxplot(fill = c("red", "blue")) +
  labs(title = "Mother's Weight by Smoking Status", y = "Weight (lbs)", x = "Smoking Status")
```

# Check normality visually
```{r}
ggplot(birthwt, aes(sample = lwt)) +
  stat_qq() + stat_qq_line() +
  facet_wrap(~smoke)
```

# Perform Wilcoxon rank sum test (non-parametric)
```{r}
wilcox_test <- wilcox.test(lwt ~ smoke, data = birthwt)
wilcox_test
```


2. Bulliforton Drug Study
a. Scientific Question and Hypothesis
H₀ (null hypothesis): The treatment with Bulliforton has no effect (success rates are equal).

H₁ (alternative hypothesis): Bulliforton affects the success rate.

Test: Chi-squared test on 2×2 contingency table.

b. Load and Analyze Primary Endpoint
# Load dataset from URL

```{r}
bull_data <- read.csv("http://www.uni-kiel.de/medinfo/lehre/medlife/bulliforton.csv")
```

# Two-way table: treatment vs primary EP
```{r}
file_path <- "C:/Users/gkakw/Desktop/Kiel/001 semester 1/004 Med Statistics/bulliforton.csv"
bull_data <- read.csv2(file_path, dec = ",")
names (bull_data)
table_primary <- table(bull_data$Treatment, bull_data$primary_EP)
table_primary
```

# Chi-squared test
```{r}
chi_primary <- chisq.test(table_primary)
chi_primary
```

c. Secondary Endpoints with Multiple Testing Correction
# Secondary EP(i)
```{r}
table_sec1 <- table(bull_data$Treatment, bull_data$sec_EP_i)
chi_sec1 <- chisq.test(table_sec1)
```

# Secondary EP(ii)
```{r}
table_sec2 <- table(bull_data$Treatment, bull_data$sec_EP_ii)
chi_sec2 <- chisq.test(table_sec2)
```

# Adjust p-values using Bonferroni
```{r}
pvals <- c(chi_sec1$p.value, chi_sec2$p.value)
p.adjust(pvals, method = "bonferroni")
```