---
title: "Exercise 6"
author: "G.K Akwaboah"
date: "`r Sys.Date()`"
output: pdf_document
---

#Part A
Scientific Question: Does smoking by the mother during pregnancy reduce the birth weight of the baby compared to the population mean of 3,200 grams?

Null Hypothesis (H₀): The mean birth weight of babies born to mothers who smoke during pregnancy is equal to the population mean, i.e., μ = 3,200 grams.

Alternative Hypothesis (Hₐ): The mean birth weight of babies born to mothers who smoke during pregnancy is less than the population mean, i.e., μ < 3,200 grams.

Significance Level (α): A significance level of α = 0.05

#Part B
The appropriate statistical test is a one-sample t-test. This test is used to compare the mean birth weight of the sample (babies of smoking mothers) to the known population mean of 3,200 grams, allowing a decision between H₀ and Hₐ based on the sample data.

Load the MASS library
```{r}
library(MASS)
```

Filter data for smoking mothers (smoke = 1)
```{r}
smoking_data <- birthwt[birthwt$smoke == 1, ]
```

Calculate mean and standard deviation

```{r}
mean_bwt <- mean(smoking_data$bwt)
sd_bwt <- sd(smoking_data$bwt)
n <- nrow(smoking_data)
```

Population mean

```{r}
mu_0 <- 3200
```

Calculate the t-test statistic

```{r}
t_stat <- (mean_bwt - mu_0) / (sd_bwt / sqrt(n))
```

Print Results

```{r}
cat("Mean birth weight:", round(mean_bwt, 2), "grams\n")
cat("Standard deviation:", round(sd_bwt, 2), "grams\n")
cat("Sample size:", n, "\n")
cat("Test statistic:", round(t_stat, 3), "\n")
```

Calculate critical value

```{r}
critical_value <- qt(0.05, df = 73, lower.tail = TRUE)
cat("Critical value:", round(critical_value, 3), "\n")
```

Perform one-sample t-test

```{r}
t_test_result <- t.test(smoking_data$bwt, mu = 3200, alternative = "less")
```

Print results
```{r}
print(t_test_result)
```

One Sample t-test
data:  smoking_data$bwt
t = -5.5859, df = 73, p-value = 7.398e-07
alternative hypothesis: true mean is less than 3200
95 percent confidence interval:
     -Inf 2934.258
sample estimates:
mean of x
  2771.933
  

